Learning Features and Abstract Actions for Computing Generalized Plans

### 2018-11-14 doc/camera-ready/paper.tex
Blai

Bonet and Geffner's generalized planning formulation assumes that the
features and abstract actions are given. In this work, we address this limitation
by  \emph{showing  how the features and abstract actions  can be learned from
the primitive predicates used to define the instances and from sampled state transitions.}
For example,  the general policy for achieving $clear(x)$ % of \citeay{bonet:ijcai2018}
is obtained using a FOND planner on an abstraction that consists of  a  boolean
feature $H$,  that tracks whether the gripper holds a block,  a numerical feature
$n(x)$ that counts the number of blocks above $x$, and two abstract actions:
one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement $n(x)$), and the other with precondition $H$ and effect $\neg H$.
We here show how to obtain such policies from STRIPS instances alone,
without having to provide the features and the abstract actions by hand.

@@

\medskip
\begin{example}
  The abstract action set $A_F = \{\bar{a}, \bar{a}'\}$, where $\bar{a} = \langle \neg H,n(x)>0; H,$ $n(x)\mminus\rangle$ %$\bar{a}=\abst{\neg H,n(x)>0}{H,n(x)\mminus}$
  and $\bar{a}'=\abst{H}{\neg H}$ , defined over the above feature set $F=\{H,n(x)\}$,
  is sound but not complete relative to $\Q_{clear}$. 
  $A_F$ captures the concrete actions that pick blocks from above $x$ and put them away,
  but not other actions like stacking a block above $x$ or picking up a block that is not above $x$.
\end{example}

@@

\noindent where  $s$ and $t$  are expanded states  $\S$ such that exactly one of them is a goal state. 
% For this, it is assumed that expanded states $s$ in $\S$ are labeled as goal or non-goal states.
%For this, it is assumed that states $s$ sampled from an instance $P$, are labeled as goal or non-goal states of $P$.
For this, it is assumed that the states in the sample are labeled as goal or non-goal states.

@@

The SAT theory $T(\S,\F)$ given by formulas \eqref{eq:d1}--\eqref{eq:goal}
has  $|\F| + m^2 (b^2 + 1)$ propositional variables, 
%has a number of propositional variables equal to $m^2 \times (b^2 + 1) + |\F|$
where $m$ is the number of expanded states in $\S$, and $b$ is their average branching factor (transitions per state).
The number of clauses is bounded by $m^2(2 + b^2 + |\F|(b^2+1))$. %$m^2 \times b^2 \times |\F|$.
%
\Omit{
This means  that  $50$ expanded states in the sample set
with an average branching factor of $10$, and a set $\F$
with $400$ candidate features, can generate up to $10^8$ clauses.
While this is a loose upper bound for the number of clauses that
follow from the implication   $D_2(s, s', t, t') \Leftharrow\ \textstyle\bigvee_f  selected(f)$
in (\label{eq:d2}) and the numbers are further reduced by symmetry consideration,
this is still a large number. ..
}
%

@@

From the point of view of expressivity, it is not clear how restrictive
is the assumption  that the  features can be obtained from a pool of  features
defined from the primitive predicates and a general grammar.
%A second limitation is the assumption that 
%the features can be obtained from a large pool of candidate  features defined 
%from the primitive predicates and a  general grammar.
%It is not fully clear whether this assumption is reasonable in general.
A relevant argument made by \citeay{bonet:ijcai2018} is that generalized problems
over domains with low width \cite{nir:ecai2012}, have compact policies
in terms of features that can be computed in polynomial time.
It is an open question, however,  whether such features are  captured
by the proposed grammar or a  suitable variation. The feature language, however, seems 
adequate for  dealing with arbitrary goals, once  \emph{goal predicates} are added
to the set of primitive predicates. Goal predicates are copies $p_G$ of the primitive predicates $p$ that
appear in the goal and which are evaluated in the goal  \cite{martin-geffner:generalized}.

Finally, the formulation can be extended to handle non-deterministic actions.
For this, it is sufficient to replace the formula \eqref{eq:bridge1}
that links $D_1$ and $D_2$ atoms  with a formula asserting that, 
if two states $s$ and $t$ are indistinguishable (i.e.\ $\neg D_1(s,t)$), then
for each action $a$ in $s$, there must be  an action $b$ in  $t$,
such that the  state transitions $(t,b,t')$ generated by $b$ in $t$, 
can't be distinguished from the transitions $(s,a,s')$ generated by $a$ in $s$. 
For this, transitions $(s,s')$ in the sample would need to be tagged with the actions
that generated them as  $(s,a,s')$.

>>>>

Bonet and Geffner's generalized planning formulation assumes that the
features and abstract actions are given. In this work, we address this limitation
by  \emph{showing  how the features and abstract actions  can be learned from
the primitive predicates used to define the instances and from sampled state transitions.}
For example,  the general policy for achieving $clear(x)$ % of \citeay{bonet:ijcai2018}
is obtained using a FOND planner on an abstraction that consists of  a  boolean
feature $H$,  that tracks whether the gripper holds a block,  a numerical feature
$n(x)$ that counts the number of blocks above $x$, and two abstract actions:
one with preconditions $\neg H$ and $n(x) > 0$, and effects $H$ and $n(x)\mminus$
(decrement of $n(x)$), and the other with precondition $H$ and effect $\neg H$.
We here show how to obtain such policies from STRIPS instances alone,
without having to provide the features and the abstract actions by hand.

@@

\medskip
\begin{example}
  The abstract action set $A_F = \{\bar{a}, \bar{a}'\}$ where $\bar{a} = \abst{\neg H,n(x)>0}{H,n(x)\mminus}$
  and $\bar{a}'=\abst{H}{\neg H}$, defined over the above feature set $F=\{H,n(x)\}$,
  is sound but not complete relative to $\Q_{clear}$.
  The actions in $A_F$ captures the concrete actions that pick blocks from above $x$ and put them away,
  but not other actions like stacking a block above $x$, or picking a block that is not above $x$.
\end{example}

@@

\noindent where  $s$ and $t$  are expanded states  in $\S$ such that exactly one of them is a goal state.
% For this, it is assumed that expanded states $s$ in $\S$ are labeled as goal or non-goal states.
%For this, it is assumed that states $s$ sampled from an instance $P$, are labeled as goal or non-goal states of $P$.
For this, it is assumed that the states in the sample are labeled as goal or non-goal states.

@@

The SAT theory $T(\S,\F)$ given by formulas \eqref{eq:d1}--\eqref{eq:goal}
has  $|\F| + m^2 (b^2 + 1)$ propositional variables,
%has a number of propositional variables equal to $m^2 \times (b^2 + 1) + |\F|$
where $m$ is the number of expanded states in $\S$, and $b$ is their average branching factor (transitions per state).
The number of clauses is bounded by $m^2(2 + b^2 + b + |\F|(b^2 + 1))$. %$m^2 \times b^2 \times |\F|$.
%
\Omit{
This means  that  $50$ expanded states in the sample set
with an average branching factor of $10$, and a set $\F$
with $400$ candidate features, can generate up to $10^8$ clauses.
While this is a loose upper bound for the number of clauses that
follow from the implication   $D_2(s, s', t, t') \Leftharrow\ \textstyle\bigvee_f  selected(f)$
in (\label{eq:d2}) and the numbers are further reduced by symmetry consideration,
this is still a large number. ..
}
%

@@

From the point of view of expressivity, it is not clear how restrictive
is the assumption  that the  features can be obtained from a pool of  features
defined from the primitive predicates and a general grammar.
%A second limitation is the assumption that
%the features can be obtained from a large pool of candidate  features defined
%from the primitive predicates and a  general grammar.
%It is not fully clear whether this assumption is reasonable in general.
A relevant argument made by \citeay{bonet:ijcai2018} is that generalized
problems over domains with bounded width \cite{nir:ecai2012} have compact
policies in terms of features $f$ whose value $\phi_f(s)$ can be computed
in polynomial time for any state $s$. It is an open question whether
such features are  captured by the proposed grammar, or a  suitable variation.
The feature language, however, seems adequate for  dealing with arbitrary goals,
once  \emph{goal predicates} are added to the set of primitive predicates.
%Goal predicates are copies $p_G$ of the primitive predicates $p$ that appear
%in the goal and which are evaluated in the goal  \cite{martin-geffner:generalized}.
Goal predicates are ``copies'' $p_G$ of the primitive predicates $p$ that
appear in the goal, and have fixed interpretation\footnote{If $p$ is
  a primitive predicate that appears in the goal, for any state $s$ and
  tuple $\bar u$ of objects, $s\vDash p_G(\bar u)$ iff $p(\bar u)$
  holds in the goal.}
\cite{martin-geffner:generalized}.

Finally, the formulation can be extended to handle non-deterministic actions.
For this, it is sufficient to replace the formula \eqref{eq:bridge1}
that links $D_1$ and $D_2$ atoms  with a formula asserting that,
if two states $s$ and $t$ are indistinguishable (i.e., $\neg D_1(s,t)$), then
for each action $a$ in $s$, there is an action $b$ in  $t$
such that the set of transitions $(t,b,t')$ generated by $b$ in $t$
cannot be distinguished from the set of transitions $(s,a,s')$ generated by $a$ in $s$.
For this, transitions $(s,s')$ in the sample would need to be tagged with the actions
that generated them as  $(s,a,s')$.
